---
title: "Exploratory data analysis"
author: "Evgenii Berdinskikh, Marina Kozyreva, Anton Tikhonenko"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = "data/pics/EDA-"
  )

library(conflicted)
library(corrplot)
library(corrr)
library(flextable)
library(forecast)
library(GGally)
library(ggpubr)
library(rstatix)
library(tidyverse)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::lag)

theme_custom <- theme_bw()+ theme(
    plot.title = element_text(size = 30, hjust = 0.5),
    plot.subtitle = element_text(size = 25, hjust = 0.5),
    strip.text = element_text(size = 20),
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 25),
    legend.title = element_text(size = 25),
    legend.text = element_text(size = 20)
  )

```

## Data import

```{r read, message=FALSE}

# International sunspot number from SIDC, Brussels (1818 - 2024)
SIDC_SILSO <- read_delim("data/raw/SIDC_SILSO.csv", delim = ";", col_names = FALSE)  %>% 
  transmute(Date = as.Date(paste0(X1, "-", X2, "-", X3)), 
            Sunspots = as.numeric(X5))

## Ambulance Data related to acute cerebrovascular accidents
brain_data <- read_csv("data/raw/brain_data.csv") %>%
   mutate(across(c(Year, Season, Month, DayOff, COVID, Sudden_Start, Dst_level), as.factor))
## Ambulance Long Data related to acute cerebrovascular accidents
brain_data_long <- read_csv("data/raw/brain_data_long.csv") %>%
    mutate(across(c(Year, Season, Month, DayOff, COVID, Sudden_Start, Dst_level, Sex, Age), as.factor))

```

## Correlation of quantitative variables

Our dataset contains a large number of quantitative variables:

```{r ggpairs, fig.height=20, fig.width=30, message=FALSE}

# Data preprocessing
brain_data_clear <- brain_data %>% 
  select( where(is.numeric) & !ends_with(c("min", "max", "var", "K_Sum")) ) %>%
  filter(complete.cases(.))

# Function for displaying correlations with a neutral color for values close to 0
color_correlation_fixed <- function(data, mapping, ...) {
  
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  cor_value <- cor(x, y, use = "complete.obs")
  
  # Defining color
  color <- if (abs(cor_value) < 0.1) {
    scales::alpha("gray50", 0.5)           ## Gray for weak correlations
  } else if (cor_value > 0) {
    scales::alpha("blue", abs(cor_value))  ## Blue for positive correlations
  } else {
    scales::alpha("red", abs(cor_value))   ## Red for negative correlations
  }
  
  # Filter for small correlation values
  cor_label <- ifelse(abs(cor_value) < 0.01, "0", sprintf("%.2f", cor_value))
  
  ggplot(data.frame()) + 
    annotate("text", x = 0.5, y = 0.5, label = cor_label, size = 10, color = color ) +
    theme_void() 
}

# Function for trend line with a golden color
golden_smooth <- function(data, mapping, ...) {
  ggplot(data, mapping) +
    geom_point(alpha = 0.4, size = 0.5, color = "black") + 
    geom_smooth(color = "gold", ...) + 
    theme_custom
}

# Plot construction
ggpairs(
  brain_data_clear, 
  progress = FALSE,
  upper = list(continuous = color_correlation_fixed), ## Highlighting correlations
  lower = list(continuous = golden_smooth) )+         ## Golden trend line
theme_custom +
  theme(
    axis.text.x = element_text(size = 10),  
    axis.text.y = element_text(size = 10) )   

```

When analyzing the network graph, three groups of features can be identified:

-   **Climatic:** temperature, pressure, wind, precipitation;
-   **Geomagnetic:** Dst index, Ap index;
-   **Solar:** F10.7 index, number of sunspots.

```{r network, fig.width=10, message=FALSE}

cor_data <- brain_data %>% 
  select( where(is.numeric) & !ends_with(c("min", "max", "Sum", "var")) ) %>%
  rename(Temperature = Temp_mean, Precipitation = H2O, 
         `F10.7 index` = F10.7, `Ap index` = Ap_mean, `Dst index` = Dst_mean) %>% 
  cor(use = "pairwise.complete.obs") 
  
corrplot(cor_data, method = 'number', type = 'lower', diag = FALSE)

network_plot(cor_data, min_cor = .1)

```

**Some features within the groups are highly correlated!**

-   A negative correlation was expected between daily average temperature and pressure (r = -0.73).
-   The F10.7 index and sunspot number show a strong correlation (r = 0.92), both reflecting solar activity in different ways.
-   The Ap and Dst indices are significantly correlated (r = -0.70), both reflecting geomagnetic field changes, though data distribution is uneven.

```{r correlation, fig.width=10, message=FALSE}

brain_data %>% 
  drop_na() %>% 
  ggplot(aes(Temp_mean, Pressure))+
  geom_jitter(alpha = 0.5, colour = "royalblue2", size = 1.5)+
  geom_smooth(method = "lm", se = FALSE, colour = "orangered2", linewidth = 1.2) +
  stat_cor(method = "pearson", label.x.npc = 0.6, label.y.npc = 0.9, size = 7)+
  labs(x = "Temperature")+
  theme_custom

brain_data %>% 
  ggplot(aes(Sunspots, F10.7))+
  geom_jitter(alpha = 0.5, colour = "royalblue2", size = 1.5)+
  geom_smooth(method = "lm", se = FALSE, colour = "orangered2", linewidth = 1.2) +
  stat_cor(method = "pearson", label.x.npc = 0, label.y.npc = 0.9, size = 7)+
  labs(x = "Sunspot Number", y = "F10.7 index")+
  theme_custom

brain_data %>% 
  ggplot(aes(Dst_mean, Ap_mean))+
  geom_jitter(alpha = 0.5, colour = "royalblue2", size = 1.5)+
  geom_smooth(method = "lm", se = FALSE, colour = "orangered2", linewidth = 1.2) +
  stat_cor(method = "pearson", label.x.npc = 0.6, label.y.npc = 0.9, size = 7)+
  labs(x = "Daily Average Dst index", y = "Daily Average Ap index")+
  theme_custom


```

Based on further analysis, the following variables will be excluded:

-   **Pressure**: does not provide full coverage of the data throughout the study period and exhibits less variability compared to Temperature.
-   **F10.7 Index**: has an incomplete data set and exhibits less variability compared to Sunspot count.
-   The **Ap and Dst** geomagnetic activity indices will not be excluded.

However, all the variables considered do not have a noticeable relationship with the number of ambulance calls:

```{r EMS, fig.height=10, fig.width=15, message=FALSE}

  brain_data %>% 
  rename(Temperature = Temp_mean, `Ap index` = Ap_mean, `Dst index` = Dst_mean) %>% 
  pivot_longer(cols = c(Temperature, Sunspots, `Ap index`, `Dst index`)) %>% 
  ggplot(aes(x = value, y = EMS))+
  geom_point()+
  geom_smooth(se = FALSE, method = "gam", colour = "goldenrod2", linewidth = 2)+
  theme_custom +
  labs(x = "", y = "Daily Emergency Calls") +  
  facet_wrap(~name, scales = "free")

```

## Data Decomposition

The number of sunspots is directly linked to the solar activity cycles, as confirmed by the graph:

```{r sunspots, fig.height=8, fig.width=12}

brain_data %>% 
  select(!c(Pressure, F10.7, Ap_mean)) %>% 
  ggplot(aes(Date, Sunspots))+
  geom_line(alpha = 0.3, colour = "royalblue2")+
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10),  se = FALSE,
              colour = "orangered4", linewidth = 2) +
  geom_vline(xintercept = as.Date("2009-01-01"), 
             colour = "orangered", linewidth = 1, linetype = "dashed")+
  geom_vline(xintercept = as.Date("2019-12-31"), 
             colour = "orangered", linewidth = 1, linetype = "dashed")+
  scale_x_date(date_breaks = "2 year", date_labels = "%Y")+
  labs(x = "", y = "Sunspot number")+
  theme_custom

```

Our data covers the complete **24th solar cycle**, spanning the period from December 2008 to December 2019.

Working with time series allows us to decompose the data to identify trends and cyclic patterns:

```{r decomposing, fig.height=15, fig.width=25}

ggarrange(

  brain_data %>% filter(Date >=  ymd("2007-01-01")) %>% pull(Sunspots) %>% 
  ts(start = 2007, frequency = 365) %>% stl(s.window = "periodic") %>% 
  autoplot() + theme_custom + ggtitle("Sunspots"), 

  brain_data %>% filter(Date >=  ymd("2007-01-01")) %>% pull(EMS) %>% 
  ts(start = 2007, frequency = 365) %>% stl(s.window = "periodic") %>% 
  autoplot() + theme_custom + ggtitle("EMS"), 

  brain_data %>% filter(Date >=  ymd("2007-01-01")) %>% pull(Temp_mean) %>% 
  ts(start = 2007, frequency = 365) %>% stl(s.window = "periodic") %>% 
  autoplot() + theme_custom + ggtitle("Temperature"), 

  nrow = 1
)

ggarrange(

  brain_data %>% filter(Date >=  ymd("2007-01-01")) %>% pull(Sunspots) %>% 
  ts(start = 2007, frequency = 365) %>% stl(s.window = "periodic") %>% 
  autoplot() + theme_custom + ggtitle("Sunspots"), 

  brain_data %>% filter(Date >=  ymd("2007-01-01")) %>% pull(Ap_mean) %>% 
  ts(start = 2007, frequency = 365) %>% stl(s.window = "periodic") %>% 
  autoplot() + theme_custom + ggtitle("Ap Index"), 

  brain_data %>% filter(Date >=  ymd("2007-01-01")) %>% pull(Dst_var) %>% 
  ts(start = 2007, frequency = 365) %>% stl(s.window = "periodic") %>% 
  autoplot() + theme_custom + ggtitle("Dst amplitude"), 
  
  nrow = 1
)

```

-   After decomposing the **emergency calls data**, a linear increase over time is observed. However, since 2020, a noticeable decrease is apparent, likely associated with the COVID-19 pandemic.
-   Climatic data (especially **temperature**), after decomposition, show a clear cyclic pattern.
-   The **daily Dst index** does not show any clear trends or cycles. However, the daily fluctuation **amplitude of the Dst index**, after decomposition, exhibits a trend similar to the dynamics of sunspot numbers. Nevertheless, since noise is the dominant component, the results should not be taken too seriously.

## Visualizing Association

### Trends: EMS & Sunspots

Solar activity shows cycles, while EMS follows a linear trend, as seen in the decomposition. However, there’s no significant difference when looking at the data by month over this period:

```{r geom_tile, fig.height=9, fig.width=18}

ggarrange(
brain_data %>%  
  filter(Date %within% interval(ymd("2007-01-01"), ymd("2019-12-31"))) %>%
  group_by(Year, Month) %>% 
  summarise(EMS_month = sum(EMS, na.rm = TRUE), .groups = "drop") %>% 
  ggplot(aes(x = Month, y = Year, fill = EMS_month)) +
  geom_tile(colour = "black") + 
  scale_fill_gradient(low = "green", high = "red") + 
  scale_x_discrete(labels = month.abb) + 
  labs(x = "", y  = "", fill = "", title = "Total Number of Emergency Calls")+
  theme_custom,

brain_data %>%  
  filter(Date %within% interval(ymd("2007-01-01"), ymd("2019-12-31"))) %>%
  group_by(Year, Month) %>% 
  summarise(Sunspots_month = sum(Sunspots, na.rm = TRUE), .groups = "drop") %>% 
  ggplot(aes(x = Month, y = Year, fill = Sunspots_month)) +
  geom_tile(colour = "black") + 
  scale_fill_gradient(low = "green", high = "red") + 
  scale_x_discrete(labels = month.abb) + 
  labs(x = "", y  = "", fill = "", title = "Total Number of Sunspots")+
  theme_custom
)

```

### Solar Cycles

When examining solar cycles over the past 200 years, sunspots can be represented as a factor variable - **Sunspots level**:

```{r Cycles, fig.height=12, fig.width=24}

Sunspots_quantile <- SIDC_SILSO %>% 
             filter(Sunspots != -1, between(Date, as.Date("1823-01-01"), as.Date("2020-01-01"))) %>% 
             pull(Sunspots) %>% quantile()
  
SIDC_SILSO %>% 
  filter(Sunspots != -1, 
         between(Date, as.Date("1823-01-01"), as.Date("2020-01-01"))) %>% 
  mutate(Sunspots_level = cut(Sunspots, 
                              breaks = c(-Inf, quantile(Sunspots)[2], quantile(Sunspots)[3], 
                                         quantile(Sunspots)[4], Inf), labels  = 
                                c("Low", "Medium", "High", "Extreme"))) %>% 
  ggplot(aes(Date, Sunspots, group = 1))+
  geom_line(aes(color = Sunspots_level), linewidth = 2)+
  geom_hline(yintercept = Sunspots_quantile[2], colour = "black", linewidth = 1, linetype = "dashed")+
  geom_hline(yintercept = Sunspots_quantile[3], colour = "black", linewidth = 1, linetype = "dashed")+
  geom_hline(yintercept = Sunspots_quantile[4], colour = "black", linewidth = 1, linetype = "dashed")+
  scale_colour_manual(values =  c("Low" = "darkorange", "Medium" = "darkorange2",  
                                  "High" = "darkorange3", "Extreme" = "darkorange4"), 
                       labels = c("Low" = "Low (0-22)", "Medium" = "Medium (23-65)", 
                                  "High" = "High (66-129)","Extreme" = "Extreme (>129)")) +
  scale_x_date( date_breaks = "20 year", date_labels = "%Y")+
  scale_y_continuous(n.breaks = 10)+
  labs(x = "", y = "Sunspot number", color = "Sunspots level", title = "Solar Cycles (1823 - 2023)")+
  theme_custom+
  annotate("text", x = as.Date("1815-01-01"), 
           y = c( (Sunspots_quantile[1] + Sunspots_quantile[2])/2, (Sunspots_quantile[2] + Sunspots_quantile[3])/2, 
                  (Sunspots_quantile[3] + Sunspots_quantile[4])/2, (Sunspots_quantile[4] + Sunspots_quantile[5])/3), 
           label = c("Q1", "Q2", "Q3", "Q4"), size = 10)

```

### EMS \~ Sunspots level

No difference in the number of emergency calls on days with different Sunspots_level is observed:

```{r level, fig.height=8, fig.width=16}

brain_data %>% 
  mutate(Sunspots_level = case_when(Sunspots > Sunspots_quantile[4] ~ "Extreme", 
                                    Sunspots > Sunspots_quantile[3] ~ "High", 
                                    Sunspots > Sunspots_quantile[2] ~ "Medium", 
                                    TRUE ~ "Low") %>% factor(fct_recode(c("Low", "Medium", "High", "Extreme")))) %>% 
 ggplot(aes(x = EMS,  y = Sunspots_level, fill = Sunspots_level )) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(
    values = c("Low" = "darkorange", "Medium" = "darkorange2", "High" = "darkorange3","Extreme" = "darkorange4"),
    labels = c("Low" = "Low (0-22)", "Medium" = "Medium (23-65)",
               "High" = "High (66-129)","Extreme" = "Extreme (>129)")) +
  labs(x = "Daily Ambulance Calls", y = "",  fill = "Sunspots level",
       title = "Relationship Between Sunspot Levels and Ambulance Calls")+
  theme_custom 



brain_data %>% 
  mutate(Sunspots_level = case_when(Sunspots > Sunspots_quantile[4] ~ "Extreme", 
                                    Sunspots > Sunspots_quantile[3] ~ "High", 
                                    Sunspots > Sunspots_quantile[2] ~ "Medium", 
                                    TRUE ~ "Low") %>% factor(fct_recode(c("Low", "Medium", "High", "Extreme")))) %>%                             
  pairwise_t_test(EMS ~ Sunspots_level, p.adjust.method = "holm") %>% 
  transmute(`Group 1` = group1, `Group 2` = group2, 
            `p-value` = p, `p.adjusted` = p.adj) %>%
  flextable() 

  
```

### Dst \~ Sunspots

At the same time, a noticeable difference in the Dst index is observed as the sunspot number changes:

```{r Dst_level, fig.height=8, fig.width=16}

brain_data %>%
  filter(Dst_level != "Severe") %>%
  mutate(Dst_level = fct_relevel(Dst_level, c("Quiet", "Weak", "Moderate", "Major", "Severe"))) %>% 
  ggplot(aes(Dst_level, Sunspots))+
  geom_jitter(alpha = 0.5, size = 1, width = 0.2) +  
  geom_boxplot(alpha = 0.5, outliers = FALSE, fill = "goldenrod2", linewidth = 1)+
  labs( x = "Geomagnetic activity level",  y = "Sunspot Number" ) +
  theme_custom

```

### EMS \~ DayOff

-   The difference between the average EMS values for weekdays and weekends is statistically significant.
-   The average EMS value is higher on weekdays (13.33) compared to weekends (11.02).

```{r DayOff, fig.height=8, fig.width=16}

mean_values <- brain_data %>%
   mutate(DayOff = fct_recode(DayOff, "Working days" = "No", "Non-working days" = "Yes")) %>% 
  group_by(DayOff) %>%
  summarise(mean_EMS = round(mean(EMS, na.rm = TRUE), 2), .groups = "drop")

brain_data %>% 
  mutate(DayOff = fct_recode(DayOff, "Working days" = "No", "Non-working days" = "Yes")) %>% 
ggplot(aes(x = DayOff, y = EMS)) +
  geom_boxplot(aes(fill = DayOff)) +
  geom_text(data = mean_values, 
            aes(x = DayOff, y = mean_EMS, label = mean_EMS), 
            color = "black", size = 8, fontface = "bold", vjust = -1) +
  scale_fill_brewer(palette = "Dark2", direction = -1)+
  scale_y_continuous(limits = c(NA, max(brain_data$EMS) + 3))+
  labs(y = "Daily Emergency Calls", x = "")+
  theme_custom +
  theme(legend.position = "none")+
  stat_pvalue_manual(t_test(data = brain_data %>%  
                              mutate(DayOff = fct_recode(DayOff, 
                                                         "Working days" = "No", 
                                                         "Non-working days" = "Yes")),  
                            EMS ~ DayOff),
                     label = "T-test, p = {p}", 
                     size = 10, 
                     y.position = max(brain_data$EMS) + 0.1)

```

### EMS \~ Season  

The distribution of EMS calls across seasons is not statistically significant (**paired t-test with Holm adjustment**) p.adjusted \> 0.05).

```{r Season, fig.height=8, fig.width=16}

brain_data %>% 
mutate(Season = fct_relevel(Season, c("Winter", "Spring", "Summer", "Autumn"))) %>% 
ggplot(aes(x = Season, y = EMS, fill = Season)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(
    values = c("Winter" = "dodgerblue2", "Spring" = "palegreen2", "Summer" = "tan2","Autumn" = "tomato2")) +
  labs(y = "Daily Emergency Calls", x = "")+
  theme_custom 

brain_data %>%
  pairwise_t_test(EMS ~ Season, p.adjust.method = "holm") %>% 
  transmute(`Group 1` = group1, `Group 2` = group2, 
            `p-value` = p, `p.adjusted` = p.adj) %>%
  flextable() 

```

### EMS \~ Age & Sex 

-   All age groups show an **increase in calls over time**, peaking around 2020.
-   **Age 25-44:** Fewer calls, with a balanced male-female distribution.
-   **Age 45-54:** More calls than 25-44, but fewer than 55+; balanced gender distribution.
-   **Age 55+:** Highest number of calls, with females surpassing males, especially in recent years.

```{r Age_Sex, fig.height=15, fig.width=20}

ggarrange(

brain_data_long %>%
  filter(between(as.numeric(as.character(Year)), 2007, 2021), Age != "55+") %>% 
  group_by(Year, Sex, Age) %>%
  summarise(EMS_total = sum(EMS, na.rm = TRUE), .groups = "drop") %>%
  mutate(EMS_plot = ifelse(Sex == "Male", -EMS_total, EMS_total)) %>% 
  ggplot(aes(x = Year, y = EMS_plot, fill = Sex)) +
  geom_bar(stat = "identity",  alpha = 0.7, colour = "black", show.legend = FALSE) +
  scale_y_continuous(labels = abs, n.breaks = 10) + 
  scale_fill_manual(
    values = c("Male" = "steelblue", "Female" = "pink"),
    labels = c("Male" = "Male", "Female" = "Female") ) +
  coord_flip() +
  labs(
    title = "Ambulance Calls Distribution by Year, Age and Gender",
    x = "", y = "", fill = "Gender") +
  theme_custom + 
  facet_wrap(. ~ Age,  scales = "free_x", 
              labeller = labeller(Age = c("25-44" = "Age: 25-44 years", 
                                          "45-54" = "Age: 45-54 years"))), 

brain_data_long %>%
  filter(between(as.numeric(as.character(Year)), 2007, 2021), Age == "55+") %>% 
  group_by(Year, Sex, Age) %>%
  summarise(EMS_total = sum(EMS, na.rm = TRUE), .groups = "drop") %>%
  mutate(EMS_plot = ifelse(Sex == "Male", -EMS_total, EMS_total)) %>% 
  ggplot(aes(x = Year, y = EMS_plot, fill = Sex)) +
  geom_bar(stat = "identity",  alpha = 0.7, colour = "black", ) +
  scale_y_continuous(labels = abs, n.breaks = 10) + 
  scale_fill_manual(
    values = c("Male" = "steelblue", "Female" = "pink"),
    labels = c("Male" = "Male", "Female" = "Female") ) +
  coord_flip() +
  labs(
    x = "", y = "Total EMS", fill = "Gender") +
  theme_custom + theme(legend.position = "inside", legend.justification = c(0.9, 0.2))+
  facet_wrap(. ~ Age,  scales = "free_x", 
              labeller = labeller(Age = c("55+" = "Age: 55+ years"))), 

nrow = 2

)
```

### Local K index

Daily Average Kp index and Daily Average K index (IRT) show a strong correlation **(r = 0.91)**. However, K index (IRT) has a smaller range and a more uniform distribution compared to Kp index.

```{r K_IRT, fig.width=10, message=FALSE}

brain_data %>% 
  drop_na(Kp_Sum, K_Sum) %>% 
  ggplot(aes(Kp_Sum, K_Sum))+
  geom_jitter(alpha = 0.5, colour = "royalblue2", size = 1.5)+
  geom_smooth(method = "lm", se = FALSE, colour = "orangered2", linewidth = 1.2) +
  stat_cor(method = "pearson", label.x.npc = 0, label.y.npc = 0.9, size = 7)+
  labs(x = "Daily Average Kp index", y = "Daily Average K index (IRT)")+
  theme_custom

brain_data %>%
  drop_na(Kp_Sum, K_Sum) %>% 
  pivot_longer(cols = c(Kp_Sum, K_Sum), names_to = "index") %>% 
  ggplot() +
  geom_histogram(aes(x = value, fill = index), 
                 binwidth = 5, colour = "black", alpha = 0.5, position = "identity", show.legend = FALSE) +
  scale_fill_manual(
    values = c("K_Sum" = "darkseagreen2", "Kp_Sum" = "steelblue2") )+
  scale_x_continuous(breaks = seq(0, 50, 10))+
  labs( x = "Index Values", y = "Observation Count", fill = "Index Type") +
  theme_custom +
  facet_wrap(~index, labeller = labeller(index = c("K_Sum" = "Daily Average K index (IRT)",
                                                   "Kp_Sum" = "Daily Average Kp index")))

```

